---
title: "Gapminder"
#author: "ID and Your Name"
date: "`r Sys.Date()`" 
output: 
  html_notebook:
    number_sections: yes
  html_document:
    df_print: paged
    number_sections: yes
---

# Gapminder and R Package `gapminder`

> Gapminder was founded by Ola Rosling, Anna Rosling Rönnlund, and Hans Rosling

-   Gapminder: <https://www.gapminder.org>

    -   Test on Top: You are probably wrong about - upgrade your worldview
    -   Bubble Chart: <https://www.gapminder.org/tools/#$chart-type=bubbles&url=v1>
    -   Dallar Street: <https://www.gapminder.org/tools/#$chart-type=bubbles&url=v1>
    -   Data: <https://www.gapminder.org/data/>

-   R Package gapminder by Jennifer Bryan

    -   Package site: <https://CRAN.R-project.org/package=gapminder>
    -   Site: <https://github.com/jennybc/gapminder>
    -   Documents: <https://www.rdocumentation.org/packages/gapminder/versions/0.3.0>

-   Package Help `?gapminder` or `gapminder` in the search window of Help

    -   The main data frame gapminder has 1704 rows and 6 variables:
        -   country: factor with 142 levels
        -   continent: factor with 5 levels
        -   year: ranges from 1952 to 2007 in increments of 5 years
        -   lifeExp: life expectancy at birth, in years
        -   pop: population
        -   gdpPercap: GDP per capita (US\$, inflation-adjusted)

# Setup

```{r setup, warning=FALSE, include=FALSE}
dir.create("data")
Sys.setenv(LANG = "en")
```

```{r packages, message=FALSE}
library(tidyverse)
library(gapminder)
library(WDI)
```

# Basics

```{r}
df <- gapminder
df
```

```{r}
glimpse(df)
```

```{r}
summary(df)
```

## Tidyverse::ggplot

## First Try - with failures

```{r}
ggplot(df, aes(x = year, y = lifeExp)) + geom_point()
```

```{r}
ggplot(df, aes(x = year, y = lifeExp)) + geom_line()
```

```{r}
ggplot(df, aes(x = year, y = lifeExp)) + geom_boxplot()
```

```{r}
typeof(pull(df, year)) # same as typeof(df$year)
```

### Box Plot

```{r}
ggplot(df, aes(x = as_factor(year), y = lifeExp)) + geom_boxplot()
```

## Applications of `dplyr`

### `filter`

```{r}
df %>% filter(country == "Afghanistan") %>%
  ggplot(aes(x = year, y = lifeExp)) + geom_line()
```

```{r}
df %>% filter(country %in% c("Afghanistan", "Japan")) %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) + geom_line()
```

```{r}
df %>% distinct(country) %>% pull()
```

```{r}
df %>% filter(country %in% c("Brazil", "Russia", "India", "China")) %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) + geom_line()
```

Russian data is missing.

### Exercises

1.  Change `lifeExp` to `pop` and `gdpPercap` and do the same.
2.  Choose ASEAN countries and do the similar investigations.

-   Brunei, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore.

3.  Choose several countries by yourself and do the similar investigations.

### `group_by` and `summarize`

Let us use the variable `continent` and summarize the data.

```{r}
df_lifeExp <- df %>% group_by(continent, year) %>% 
  summarize(mean_lifeExp = mean(lifeExp), median_lifeExp = median(lifeExp), max_lifeExp = max(lifeExp), min_lifeExp = min(lifeExp), .groups = "keep")
```

```{r}
df_lifeExp
```

```{r}
df %>% filter(year %in% c(1952, 1987, 2007)) %>%
  ggplot(aes(x=as_factor(year), y = lifeExp, fill = continent)) +
  geom_boxplot()
```

```{r}
df_lifeExp %>% ggplot(aes(x = year, y = mean_lifeExp, color = continent)) +
  geom_line()
```

```{r}
df_lifeExp %>% ggplot(aes(x = year, y = mean_lifeExp, color = continent, linetype = continent)) +
  geom_line()
```

```{r}
df_lifeExp %>% ggplot() +
  geom_line(aes(x = year, y = mean_lifeExp, color = continent)) + 
  geom_line(aes(x = year, y = median_lifeExp, linetype = continent))
```

# Examples in Gapminer Package Help

```{r}
str(gapminder)
```

```{r}
head(gapminder)
```

```{r}
summary(gapminder)
```

```{r}
table(gapminder$continent)
```

```{r}
aggregate(lifeExp ~ continent, gapminder, median)
```

```{r}
plot(lifeExp ~ year, gapminder, subset = country == "Cambodia", type = "b")
```

```{r}
plot(lifeExp ~ gdpPercap, gapminder, subset = year == 2007, log = "x")
```

```{r}
gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(lifeExp = median(lifeExp))
```

How many unique countries does the data contain, by continent?

```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(n_obs = n(), n_countries = n_distinct(country))
```

By continent, which country experienced the sharpest 5-year drop in life expectancy and what was the drop?

```{r}
gapminder %>%
  group_by(continent, country) %>%
  select(country, year, continent, lifeExp) %>%
  mutate(le_delta = lifeExp - lag(lifeExp)) %>%
  summarize(worst_le_delta = min(le_delta, na.rm = TRUE)) %>%
  filter(min_rank(worst_le_delta) < 2) %>%
  arrange(worst_le_delta)
```

# Additional Exploration Using Tidyverse

# World Development Indicators (WDI)

-   SP.DYN.LE00.IN: Life expectancy at birth, total (years)
-   NY.GDP.PCAP.KD: GDP per capita (constant 2015 US\$)
-   SP.POP.TOTL: Population, total

```{r cash=TRUE}
df_wdi <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD")
)
```

```{r}
df_wdi
```

```{r cash=TRUE}
df_wdi_extra <- WDI(
  country = "all", 
  indicator = c(lifeExp = "SP.DYN.LE00.IN", pop = "SP.POP.TOTL", gdpPercap = "NY.GDP.PCAP.KD"), 
  extra = TRUE
)
```

```{r}
df_wdi_extra
```

```{r}
WDIsearch(string="gdp per capita", field="name", short=FALSE, cache=NULL)
```

```{r}
WDIsearch(string="life expectancy", field="name", short=FALSE, cache=NULL)
```

```{r}
WDIsearch(string="population, total", field="name", short=FALSE, cache=NULL)
```

There are a lot of missing data. Can you do similar investigations as we did using `gapminder` package?

```{r}
df_wdi %>% distinct(country) %>% pull()
```

```{r}
df_wdi %>% filter(country %in% c("Brazil", "Russian Federation", "India", "China")) %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) + geom_line()

```

```{r}
df_wdi_lifeExp <- df_wdi_extra %>% group_by(region, year) %>% 
  summarize(mean_lifeExp = mean(lifeExp), median_lifeExp = median(lifeExp), max_lifeExp = max(lifeExp), min_lifeExp = min(lifeExp), .groups = "keep")
```

```{r}
df_wdi_lifeExp
```

```{r}
df_wdi_extra %>% filter(year %in% c(1952, 1987, 2007)) %>%
  ggplot(aes(x=as_factor(year), y = lifeExp, fill = region)) +
  geom_boxplot()
```

Only with not NA values.

```{r}
df_wdi_extra %>% filter(year %in% c(1952, 1987, 2007), !is.na(lifeExp), !is.na(region)) %>%
  ggplot(aes(x=as_factor(year), y = lifeExp, fill = region)) +
  geom_boxplot()
```

```{r}
df_wdi_lifeExp %>% ggplot(aes(x = year, y = mean_lifeExp, color = region, linetype = region)) +
  geom_line()
```

```{r}
df_wdi_lifeExp %>% filter(!is.na(mean_lifeExp), !is.na(median_lifeExp)) %>%
                            ggplot() +
  geom_line(aes(x = year, y = mean_lifeExp, color = region)) + 
  geom_line(aes(x = year, y = median_lifeExp, linetype = region))
```
