---
title: "Japan as a part of the world"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: haddock
    theme: cosmo
    number_sections: yes
  pdf_document:
    toc: yes
---

> In this paper, we look at the data of the world and Japan using the World Development Indicators.

# World Development Indicators

## World Bank: WDI - World Development Indicaters

* World Bank: https://www.worldbank.org
* [Who we are](https://www.worldbank.org/en/who-we-are):
  - To end extreme poverty: By reducing the share of the global population that lives in extreme poverty to 3 percent by 2030.
  - To promote shared prosperity: By increasing the incomes of the poorest 40 percent of people in every country. 
* World Bank Open Data: https://data.worldbank.org
  - Data Bank, World Development Indicators, etc.
* [World Development Indicators (WDI)](https://datatopics.worldbank.org/world-development-indicators/) : the World Bank’s premier compilation of cross-country comparable data on development; 1400 time series indicators
  - Themes: Poverty and Inequality, People, Environment, Economy, States and Markets, Global Links
  - Open Data & DataBank: Explore data, Query database
  - Bulk Download: Excel, CSV
  - API Documentation
  
---

## R Package [WDI](https://CRAN.R-project.org/package=WDI)

* [WDI](https://CRAN.R-project.org/package=WDI): World Development Indicators and Other World Bank Data
* Search and download data from over 40 databases hosted by the World Bank, including the World Development Indicators ('WDI'), International Debt Statistics, Doing Business, Human Capital Index, and Sub-national Poverty indicators.
* Version: 2.7.4
* Materials:	[README](https://cran.r-project.org/web/packages/WDI/readme/README.html)   - _usage_
  - [NEWS](https://cran.r-project.org/web/packages/WDI/news/news.html) - _version history_
* Published: 2021-04-06
* README: https://cran.r-project.org/web/packages/WDI/readme/README.html
* Reference manual:	[WDI.pdf](https://cran.r-project.org/web/packages/WDI/WDI.pdf)

# Set up

```{r}
library(tidyverse)
library(WDI)
library(readxl)
```
```{r cache = TRUE}
wdi_cache <- WDIcache()
```


# World Development Indicators (WDI)

## Setup


## Indicators Used in Assignment Three

* NY.GDP.PCAP.KD: GDP per capita (constant 2015 US$)
* NY.GDP.MKTP.CD: GDP (current US$)
* NY.GDP.PCAP.KD.ZG: GDP per capita growth (annual%)
* SP.POP.TOTL: Total Population
* IP.JRN.ARTC.SC: Scientific and technical journal articles
* FP.CPI.TOTL.ZG: Inflation, consumer prices (% annual growth) 
* DT.ODA.ODAT.PC.ZS: Net ODA received per capita (current US$)
* BX.KLT.DINV.CD.WD: Foreign direct investment, net inflows (BoP, current US$)
* MS.MIL.XPND.GD.ZS: Military expenditure (% of GDP)
* MS.MIL.TOTL.P1: Total Armed Forces Personnel
* SI.POV.MDIM: Multidimensional poverty headcount ratio (% of total population)
* SI.POV.NAHC: Poverty headcount ratio at national poverty lines (% of population) 
* SI.POV.GINI: Gini index
* NY.GNS.ICTR.ZS: Gross savings (% of GDP)
* SL.UEM.TOTL.ZS: Unemployment, total (% of total labor force) (modeled ILO estimate)
* SP.URB.TOTL.IN.ZS (Urban population (% of total population))
* AG.LND.AGRI.ZS: Agricultural land (% of land area)
* SE.XPD.TOTL.GD.ZS: Government expenditure on education, total (% of GDP)
* SG.GEN.PARL.ZS: Proportion of seats held by women in national parliaments (%)

## WDIsearch


```{r}
x <- "SI.POV.MDIM"
wdis <- function(x) WDIsearch(string = x, field = "indicator", short = FALSE, cache = wdi_cache)
wdis(x)
```


```{r}
library(broom)
wdi_study <- c("NY.GDP.PCAP.KD", "NY.GDP.MKTP.CD", "NY.GDP.PCAP.KD.ZG", "SP.POP.TOTL", "IP.JRN.ARTC.SC", "FP.CPI.TOTL.ZG","DT.ODA.ODAT.PC.ZS", "BX.KLT.DINV.CD.WD","MS.MIL.XPND.GD.ZS", "MS.MIL.TOTL.P1", "SI.POV.MDIM", "SI.POV.NAHC", "SI.POV.GINI", "NY.GNS.ICTR.ZS", "SL.UEM.TOTL.ZS","SP.URB.TOTL.IN.ZS", "AG.LND.AGRI.ZS","SE.XPD.TOTL.GD.ZS", "SG.GEN.PARL.ZS")

wdis <- function(x) WDIsearch(string = x, field = "indicator", short = FALSE, cache = wdi_cache) %>% head(1)

sapply(wdi_study, wdis, simplify=TRUE) %>% as_tibble() %>% unnest()
```

```{r}
wdis("NY.GDP.PCAP.KD")
```


```
WDIsearch(string = "gdp", field = "name", short = TRUE, cache = NULL)
```



```{r}
WDIsearch("NY.GNS.ICTR.ZS", "indicator")
```
* The simplest format 2.

```{r}
WDIsearch("Literacy rate", "name")
```

* Detailed with short = FALSE

```{r}
WDIsearch("IP.JRN.ARTC.SC", "indicator", FALSE)
```
* You can use `wdi_cache` downloaded above to use the updated meta data.

```{r}
WDIsearch("Government expenditure on education", "name", FALSE, wdi_cache)
```


## Importing Data Using WDI

* The shortest form. The following is same as:
`WDI(country = "all", indicator = "NY.GDP.PCAP.KD", start = 1960, end = NULL, extra = FALSE, cache = NULL, latest = NULL, language = "en)`

```{r}
WDI(indicator = "NY.GDP.PCAP.KD")
```

* In order to use the downloaded data, we need to assign a name to it. It can be `df` if you download and use only one data, but it is better to assign the data to be a more descriptive name.

```{r cache = TRUE}
df_gdppcap <- WDI(indicator = "NY.GDP.PCAP.KD")
df_gdppcap
```

* If you want to use extra information such as `region`, `income`, `lending`, use `extra = TRUE`.

```{r cache = TRUE}
df_gdppcap_extra <- WDI(indicator = "NY.GDP.PCAP.KD", extra = TRUE)
df_gdppcap_extra
```

## Writing and reading a csv file

Since the data is generally huge, it is better to use a data saved in your computer.
Before saving it check whether you have `data` folder (or directory) in you project folder. Use the `Files` tab of the right bottom pane. You should be able to see the project icon with Rproj at the end, your R Notebook file you are editing, and the `data` folder. Since `.csv` is automatically added, `write_csv(gdppcap_extra, "data/gdppcap_extra")` does the same as below.

```{r eval = FALSE}
write_csv(gdppcap_extra, "data/gdppcap_extra.csv")
```

To read the data, run the next code chunk. You do not have to run the code `df_gdppcap_extra <- WDI(indicator = "NY.GDP.PCAP.KD", extra = TRUE)` again.

```{r eval = FALSE}
gdppcap <- read_csv("data/gdppcap_extra.csv")
```

"data/gdpcap_extra.csv" is the way to express the name of the data `gdpcap_extra.csv` in csv format in the data folder. 

## `country` argument

If you want to import data for several countries, you can use `iso2c` codes of the countries.

```{r cache = TRUE}
ASEAN <- c("BN", "ID", "KH", "LA", "MM", "MY", "PH", "SG")
df_gdppcap_asean <- WDI(ASEAN, "NY.GDP.PCAP.KD")
df_gdppcap_asean %>% distinct(country) %>% pull()
```


```{r}
wdi_cache$country %>% filter(iso2c %in% ASEAN) %>% distinct(country) %>% pull()
```


You can also use `wdi_cache$country`.

```{r}
wdi_cache$country
```

```{r}
wdi_cache$country %>% filter(iso2c %in% ASEAN) %>% pull(country)
```


You can find iso2c codes from the downloaded data, or using wdi_cache$country

```{r}
wdi_cache$country %>% 
  filter(country %in% c("Brunei Darussalam", "Indonesia", "Cambodia", "Lao PDR", "Myanmar", "Malaysia", "Philippines", "Singapore")) %>%
  pull(iso2c)
```

You can also find countries in some category.

```{r}
wdi_cache$country %>% 
  filter(region == "South Asia") %>%
  pull(country)
```


```{r}
wdi_cache$country %>% 
  filter(income == "Lower middle income") %>%
  pull(iso2c)
```

The following shows a list of indicators

```{r}
wdi_cache$series
```

The following is same as `WDIsearch("gdp per cap", "name")`. See WDIsearch help.

```{r}
wdi_cache$series %>% filter(grepl("gdp per cap", name, ignore.case = TRUE))
```


```{r cache = TRUE}
wdi_cache$series %>% filter(indicator == "MS.MIL.XPND.GD.ZS") %>% pull(name)
df_milxpnd <- WDI(indicator = c(milxpnd = "MS.MIL.XPND.GD.ZS"), extra = TRUE) %>%
  drop_na(milxpnd)
```


Default of the number of bins is 30. Here I chose binwidth = 0.5 (%).

```{r}
df_milxpnd %>% filter(income != "Aggregates", year == 2021) %>% 
  ggplot(aes(milxpnd, fill = income)) + geom_histogram(color = "black", binwidth = 0.5) + 
  labs(title = "Military expenditure (% of GDP)")
```

```{r}
df_milxpnd %>% filter(income != "Aggregates", income != "Not classified", year == 2021) %>% 
  ggplot(aes(milxpnd, fill = income)) + geom_density(binwidth = 0.5, alpha = 0.3) + 
  labs(title = "Military expenditure (% of GDP)")
```


```{r cache = TRUE}
wdi_cache$series %>% filter(indicator == "SL.UEM.TOTL.ZS") %>% pull(name)
df_ur <- WDI(indicator = c(ur = "SL.UEM.TOTL.ZS"), 
             extra = TRUE, cache = wdi_cache) %>% drop_na(ur)
```


```{r}
df_ur %>% filter(income == "Aggregates") %>% filter(grepl('income', country)) %>% 
  filter(year >= 2018) %>% ggplot() + geom_line(aes(x = year, y = ur, color = country))
```


```{r cache = TRUE}
wdi_cache$series %>% filter(indicator == "FP.CPI.TOTL.ZG") %>% pull(name)
df_infl <- WDI(country = c("VN","CN","JP"), 
               indicator = c(inf = "FP.CPI.TOTL.ZG"), extra = TRUE) %>% drop_na(inf)
```

```{r}
df_infl_s <- df_infl %>% filter(year %in% c(2000, 2005, 2010, 2015, 2020), 
                   country %in% c("Japan", "Vietnam", "China"))
df_infl_s
```
```{r}
df_infl_s %>%
  ggplot(aes(x = year, y = inf, color= country)) + 
  geom_line() + geom_point() + 
  geom_text(aes(label = round(inf,2)), nudge_y = 0.8) + 
labs(title = "Inflation of Japan, Vietnam and China from 2000 to 2020", 
     x = "", y = "Inflation, consumer prices (annual %)")
```

List the name of countries of ASEAN and BRICs using `wdi_cache$country`.

```{r}
asean <- c("Brunei Darussalam", "Cambodia", "Lao PDR", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore")
brics <- c("Brazil", "Russian Federation", "India", "China")
```

Find the `iso2c` of the countries using `wdi_cache$country`.

```{r}
wdi_cache$country %>% 
  filter(country %in% 
           c("Brunei Darussalam", "Cambodia", "Lao PDR", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore",
           "Brazil", "Russian Federation", "India", "China")) %>% 
  pull(iso2c)
```

Separate the iso3c's of the countries with commas and read data using `WDI`.

```{r cache=TRUE}
wdi_gdp <- WDI(
  country = c("BR", "BN", "CN", "ID", "IN", "KH", "LA", "MM", "MY", "PH", "RU", "SG"),
  indicator = c(gdp = "NY.GDP.MKTP.KD", gdpPercap = "NY.GDP.PCAP.KD"),
  start = 1960, extra = TRUE, cache = wdi_cache)
```


```{r}
wdi_gdp %>% filter(country %in% asean) %>% drop_na(gdp, gdpPercap) %>% summary() 
```





