---
title: "Maps Test"
author: "Hiroshi Suzuki"
output:
  html_notebook: default
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```

# `geom_map`

```
map_data(map, region = ".", exact = FALSE, ...)
```

```{r}
map_data('world') %>% as_tibble()
```

```{r}
world0 <- map_data('world')
world0 %>%
  ggplot() +
  geom_map(aes(long, lat, map_id = region), map = world0, fill = "gray", col = "black", size = 0.1) 
```

```{r}
library(maps)
iso.expand("JP")
iso3166
```

```{r}
library(WDI)
wdi_cache <- WDIcache()
wdi_cache
```

```{r}
world0 <- map_data('world')
wdi_income <- wdi_cache$country %>% select(iso2c, income)
world_income <- world0 %>% mutate(iso2c = iso.alpha(region, n=2)) %>%
  left_join(wdi_income, by = "iso2c") 
world_income
```

```{r}
world_income %>% 
  mutate(income_level = factor(income, levels = c("High income", "Upper middle income", "Lower middle income", "Low income", "Not classified", NA))) %>%
  ggplot() +
  geom_map(aes(long, lat, map_id = region, fill = income_level), map = world_income, col = "black", size = 0.1) 
```

```{r}
world0 %>% left_join(iso3166, by = c("region" = "mapname")) 
```

```{r}
wdi_cache$country
```


```{r}
eastasia <- map_data('world', region = c("Japan", "South Korea", "North Korea", "China", "Mongolia", "Taiwan")) %>%
  mutate(iso2c = iso.alpha(region, n=2))
eastasia %>%
  ggplot() +
  geom_map(aes(long, lat, map_id = region), map = eastasia, fill = "gray", col = "black", size = 0.1) 
```

```{r}
eastasia <- map_data('world', region = c("Japan", "South Korea", "North Korea", "China", "Mongolia", "Taiwan")) %>% mutate(iso2c = iso.alpha(region, n=2))
eastasia %>% left_join(wdi_income, by = "iso2c") %>% 
  mutate(income_level = factor(income, levels = c("High income", "Upper middle income", "Lower middle income", "Low income", "Not classified", NA))) %>%
  ggplot() +
  geom_map(aes(long, lat, map_id = region, fill = income_level), map = eastasia, col = "black", size = 0.1) 
```

```{r}
china <- map_data('world', region = c("China"))
china %>%
  ggplot() +
  geom_map(aes(long, lat, map_id = region, fill = subregion), map = china, col = "black", size = 0.1) 
```

```{r}
ea <- map_data('world', region = c("Japan", "South Korea", "North Korea", "China", "Mongolia", "Taiwan"))
ea %>%
  ggplot() +
  geom_map(aes(long, lat, map_id = region, fill = region), map = ea, col = "black", size = 0.1) +
  labs(x = "", y = "") + 
  theme(legend.position="none", panel.grid = element_blank(), 
        axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.text.y=element_blank(), axis.ticks.y=element_blank())
```

```{r}
map_data('world') %>% ggplot(aes(long, lat)) + geom_polygon(aes(group = group, fill = region), col = "black", linewidth = 0.1) + theme(legend.position="none")
```

```{r}
map_data('world2') %>% ggplot(aes(long, lat)) + geom_polygon(aes(group = group, fill = region), col = "black", linewidth = 0.1) + theme(legend.position="none")
```

```{r}
world1 <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE))
ggplot() + geom_sf(data = world1)
```


```{r}
world_sf <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE)) %>% 
  mutate(iso2c = iso.alpha(ID, n=2)) %>% 
  left_join(wdi_income, by = "iso2c") 
world_sf %>% as_tibble()
```

```{r}
world_sf %>% 
  mutate(income_level = factor(income, levels = c("High income", "Upper middle income", "Lower middle income", "Low income", "Not classified", NA))) %>%
  ggplot() +
  geom_sf(aes(fill = income_level)) 
```

```{r}
world2 <- sf::st_transform(
  world1,
  "+proj=laea +y_0=0 +lon_0=155 +lat_0=-90 +ellps=WGS84 +no_defs"
)
ggplot() + geom_sf(data = world2)
```

### Volcano Data

https://datavizpyr.com/how-to-make-world-map-with-ggplot2-in-r/

```{r}
volcano <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv")
volcano
```

```{r}
ggplot() +
  geom_map(
    data = world0, map = world0,
    aes(long, lat, map_id = region),
    color = "white", fill = "lightgray", size = 0.1
  ) +
  geom_point(
    data = volcano,
    aes(longitude, latitude, color = primary_volcano_type),
    alpha = 0.7
  ) 
```


## `annotation_map`

```{r}
df <- data.frame(
  name = c("Charlotte", "Raleigh", "Greensboro"),
  lat = c(35.227, 35.772, 36.073),
  long = c(-80.843, -78.639, -79.792)
)

p <- ggplot(df, aes(x = long, y = lat)) +
  annotation_map(
    map_data("state"),
    fill = "antiquewhite", colour = "darkgrey"
  ) +
  geom_point(color = "blue") +
  geom_text(
    aes(label = name),
    hjust = 1.105, vjust = 1.05, color = "blue"
  )

p + xlim(-84, -76) + ylim(34, 37.2)
```

```{r}
p +
  coord_sf(
    crs = sf::st_crs(3347),
    default_crs = sf::st_crs(4326),  # data is provided as long-lat
    xlim = c(-84, -76),
    ylim = c(34, 37.2)
  )
```

```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
p +
  geom_sf(
    data = nc, inherit.aes = FALSE,
    fill = NA, color = "black", linewidth = 0.1
  ) +
  coord_sf(crs = sf::st_crs(3347), default_crs = sf::st_crs(4326))
```

# `geom_sf`

See also, `coord_sf`, `geom_sf_label`, `geom_sf_text`, `stat_sf`.

> This set of geom, stat, and coord are used to visualise simple feature (sf) objects. For simple plots, you will only need geom_sf() as it uses stat_sf() and adds coord_sf() for you. geom_sf() is an unusual geom because it will draw different geometric objects depending on what simple features are present in the data: you can get points, lines, or polygons. For text and labels, you can use geom_sf_text() and geom_sf_label().

```
geom_sf(
  mapping = aes(),
  data = NULL,
  stat = "sf",
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE,
  ...
)
```

```{r}
# map_data {ggplot2}
# Easily turn data from the maps package into a data frame suitable for plotting with ggplot2.
world_map <- map_data("world")
```


```{r}
library(maps)
world1 <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE))
ggplot() + geom_sf(data = world1)
```

```{r}
world_map <- map('world', plot = FALSE)
glimpse(world_map)
```

# `geom_polygon`

```{r}
# When using geom_polygon, you will typically need two data frames:
# one contains the coordinates of each polygon (positions),  and the
# other the values associated with each polygon (values).  An id
# variable links the two together

ids <- factor(c("1.1", "2.1", "1.2", "2.2", "1.3", "2.3"))

values <- data.frame(
  id = ids,
  value = c(3, 3.1, 3.1, 3.2, 3.15, 3.5)
)

positions <- data.frame(
  id = rep(ids, each = 4),
  x = c(2, 1, 1.1, 2.2, 1, 0, 0.3, 1.1, 2.2, 1.1, 1.2, 2.5, 1.1, 0.3,
  0.5, 1.2, 2.5, 1.2, 1.3, 2.7, 1.2, 0.5, 0.6, 1.3),
  y = c(-0.5, 0, 1, 0.5, 0, 0.5, 1.5, 1, 0.5, 1, 2.1, 1.7, 1, 1.5,
  2.2, 2.1, 1.7, 2.1, 3.2, 2.8, 2.1, 2.2, 3.3, 3.2)
)

# Currently we need to manually merge the two together
datapoly <- merge(values, positions, by = c("id"))

p <- ggplot(datapoly, aes(x = x, y = y)) +
  geom_polygon(aes(fill = value, group = id))
p
```

```{r}
# Which seems like a lot of work, but then it's easy to add on
# other features in this coordinate system, e.g.:

set.seed(1)
stream <- data.frame(
  x = cumsum(runif(50, max = 0.1)),
  y = cumsum(runif(50,max = 0.1))
)

p + geom_line(data = stream, colour = "grey30", linewidth = 5)
```

```{r}
# And if the positions are in longitude and latitude, you can use
# coord_map to produce different map projections.

  # As of R version 3.6 geom_polygon() supports polygons with holes
  # Use the subgroup aesthetic to differentiate holes from the main polygon

  holes <- do.call(rbind, lapply(split(datapoly, datapoly$id), function(df) {
    df$x <- df$x + 0.5 * (mean(df$x) - df$x)
    df$y <- df$y + 0.5 * (mean(df$y) - df$y)
    df
  }))
  datapoly$subid <- 1L
  holes$subid <- 2L
  datapoly <- rbind(datapoly, holes)

  p <- ggplot(datapoly, aes(x = x, y = y)) +
    geom_polygon(aes(fill = value, group = id, subgroup = subid))
  p
```

# Natural Earth Data 

https://www.naturalearthdata.com

Get natural earth world country polygons

Manual: https://cran.r-project.org/web/packages/rnaturalearth/rnaturalearth.pdf

## `ne_countries`

```
ne_countries(
  scale = 110,
  type = "countries",
  continent = NULL,
  country = NULL,
  geounit = NULL,
  sovereignty = NULL,
  returnclass = c("sp", "sf")
)
```

### Arguments

* scale: scale of map to return, one of 110, 50, 10 or 'small', 'medium', 'large'
* type: country type, one of 'countries', 'map_units', 'sovereignty', 'tiny_countries'
* continent: a character vector of continent names to get countries from.
* country: a character vector of country names.
* geounit: a character vector of geounit names.
* sovereignty: a character vector of sovereignty names.
* returnclass: 'sp' default or 'sf' for Simple Features

```{r}
ne_countries(scale = "large", returnclass = "sf") %>%
  ggplot() +   geom_sf()
```

```{r}
ne_countries(scale = "small", returnclass = "sf") %>%
  ggplot() +   geom_sf()
```

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
class(world)
```

```{r}
world %>% as_tibble()
```
```{r}
world %>% colnames()
```

```{r}
world %>% ggplot() +   geom_sf()
```

```{r}
world %>% ggplot() +   geom_sf(aes(fill = income_grp))
```

```{r}
world %>% filter(subregion == "South-Eastern Asia") %>%
  ggplot() +   geom_sf(aes(fill = income_grp))
```

### `type` argument

```{r}
ne_countries(type = "countries", country = "Japan", scale = "medium", returnclass = "sf") %>%
  ggplot() + geom_sf()
```

```{r}
ne_countries(type = "map_units", country = "Japan", scale = "medium", returnclass = "sf") %>%
  ggplot() + geom_sf()
```

```{r}
library(sp)
  sf_world <- ne_countries(returnclass = "sf")
  plot(sf_world)
spdf_tiny_countries <- ne_countries(type = "tiny_countries", scale = 50)
  plot(spdf_tiny_countries)
```


## `ne_states`

Get natural earth world state (admin level 1) polygons

**Description**: returns state polygons (administrative level 1) for specified countries

### Usage

```
ne_states(
  country = NULL,
  geounit = NULL,
  iso_a2 = NULL,
  spdf = NULL,
  returnclass = c("sp", "sf")
)
```

### Arguments

* country: a character vector of country names.

* geounit: a character vector of geounit names.

* iso_a2: a character vector of iso_a2 country codes

* spdf: an optional alternative states map

* returnclass: 'sp' default or 'sf' for Simple Features

### Value

* SpatialPolygons DataFrame or sf

```{r}
world_admin1 <- ne_states(returnclass = "sf")
```

```{r}
class(world_admin1)
```

```{r}
world_admin1 %>% colnames()
```

```{r}
world_admin1 |> as_tibble()
```

```{r}
world_admin1 |> glimpse()
```

```{r}
world_admin1 %>% as_tibble() %>% 
  filter(iso_a2 != "-1") %>% arrange(admin) %>%
  distinct(iso_a2, admin)
```

```{r}
world_admin1 %>% pull(FCLASS_JP) %>% unique()
```

```{r}
japan <- world_admin1[world_admin1$admin == "Japan",]
```

```{r}
japan |> as_tibble()
```

```{r}
ggplot() +   geom_sf(data = japan) +   theme_minimal()
```

```{r}
country <- "Japan"
world_admin1 %>% filter(admin == country) %>% 
  ggplot() + geom_sf()
```

```{r}
iso2 <- "ID"
world_admin1 %>% filter(iso_a2 == iso2) %>% 
  ggplot() +   geom_sf()
```

```{r}
iso2s <- c("IN","PK","BD","LK")
world_admin1 %>% filter(iso_a2 %in% iso2s) %>% 
  ggplot() +   geom_sf(aes(fill = admin))
```

```{r}
regions <- c("Tohoku")
world_admin1 %>% filter(region %in% regions) %>% 
  ggplot() +   geom_sf()
```

```{r}
world_admin1 %>% filter(iso_a2 == "JP") %>%
  ggplot() +   geom_sf(aes(fill = region))
```

```{r}
world_admin1 %>% filter(iso_a2 == "JP") %>%
  ggplot() +   geom_sf(aes(fill = name_local))
```

```{r}
world_admin1 %>% filter(iso_a2 == "JP") %>%
  ggplot() +   geom_sf(aes(fill = name_local)) +
  theme(legend.position = "none")
```

```{r}
world_admin1 %>% arrange(admin) %>% pull(admin) %>% unique()
```


## `ne_download`

download data from Natural Earth and (optionally) read into R

```{r}
ne_download(scale = 110, type = "countries", returnclass = "sf") %>%
  ggplot() + geom_sf()
```

```{r}
# for raster, here an example with Manual Shaded Relief (MSR) download & load

rst <- ne_download(scale = 50, type = "MSR_50M", category = "raster")

# load after having downloaded
rst <- ne_load(
  scale = 50, type = "MSR_50M", category = "raster", destdir =
    getwd()
)

# plot
library(terra)
terra::plot(rst)
# end dontrun
```

# `geodata` Package

* URL: https://gadm.org
* `geodata`: Download Geographic Data
* Functions for downloading of geographic data for use in spatial analysis and mapping. The package facilitates access to climate, crops, elevation, land use, soil, species occurrence, accessibility, administrative boundaries and other data.
* Package Link:  https://CRAN.R-project.org/package=geodata
* Manual: https://cran.r-project.org/web/packages/geodata/geodata.pdf

## `gadm` Administrative boundaries

* Get administrative boundaries for any country in the world. Data are read from files that are down- loaded if necessary.
* Usage: `gadm(country, level=1, path, version="latest", resolution=1, ...)`
* Arguments
    - country: character. Three-letter ISO code or full country name. If you provide multiple names they are all downloaded and rbind-ed together
    - level: numeric. The level of administrative subdivision requested. (starting with 0 for country, then 1 for the first level of subdivision)
    - path: character. Path for storing the downloaded data. See geodata_path
    - version: character. Either "latest" or GADM version number (can be "3.6", "4.0" or "4.1")
    - resolution: integer indicating the level of detail. Only for version 4.1. It should be either 1 (high) or 2 (low)

```{r}
library(geodata)
```

```{r}
world(resolution=5, level=0, path = "./data")
```
```{r}
world5 <- readRDS("./data/gadm/gadm36_adm0_r5_pk.rds")
world5 %>% as_tibble() %>% glimpse()
```

```{r}
world5 %>% st_as_sf() %>% ggplot() + geom_sf()
```

```{r}
world5[world5$GID_0 == "JPN",] %>% st_as_sf() %>% ggplot() + geom_sf()
```

```{r}
world5 %>% 
  st_as_sf() %>% filter(GID_0 == "JPN") %>% 
  ggplot() + geom_sf()
```

```{r}
world(resolution=1, level=0, path = "./data") %>%
  st_as_sf() %>% ggplot() + geom_sf()
```
```{r}
world(path = "./data") %>%
  st_as_sf() %>% ggplot() + geom_sf()
```

```{r}
world(resolution=1, level=0, path = "./data") %>%
  st_as_sf() %>% filter(NAME_0 == "Japan") %>%
  ggplot() + geom_sf()
```

```{r}
world(resolution=1, level=0, path = "./data") %>%
  st_as_sf() %>% filter(NAME_0 %in% c("India","Pakistan", "Bangladesh" , "Sri Lanka")) %>%
  ggplot() + geom_sf(aes(fill = GID_0))
```

```{r}
world5_df <- st_as_sf(world5) %>% 
  st_set_crs("+proj=longlat +datum=WGS84") %>% 
  st_transform(., "+proj=robin")

ggplot() +
  geom_sf(data = world5_df)
```

See https://r.geocompx.org/reproj-geo-data.html

```{r}
world5_df <- st_as_sf(world5) %>% 
  st_set_crs("+proj=longlat +datum=WGS84") %>%
  st_transform("+proj=robin")

ggplot() +
  geom_sf(data = world5_df)
```


```{r}
world5_df <- st_as_sf(world5) %>% 
  st_set_crs("+proj=longlat +datum=WGS84") %>%
  st_transform("+proj=moll")

ggplot() +
  geom_sf(data = world5_df)
```
Error

```{r}
st_as_sf(world5) %>% 
  st_set_crs("+proj=longlat +datum=WGS84") %>%
  st_transform(crs = "+proj=wintri") %>% 
ggplot() +
  geom_sf()
```

### `gadm`

```{r}
gadm0 <- gadm("JPN", level = 0, path = "./data/")
```
```{r}
gadm0 %>% st_as_sf() %>% 
  ggplot() + geom_sf()
```

```{r}
gadm1 <- gadm("JPN", level = 1, path = "./data/") %>% st_as_sf()
gadm1 %>% glimpse()
```

```{r}
gadm1 %>% 
  ggplot() + geom_sf(aes(fill = NAME_1)) + theme(legend.position = "none")
```

```{r}
gadm1 %>% st_set_crs("+proj=longlat +datum=WGS84") %>%
  st_transform("+proj=robin") %>%
  ggplot() + geom_sf(aes(fill = NAME_1)) + theme(legend.position = "none")
```


```{r}
gadm2 <- gadm("JPN", level = 2, path = "./data/") %>% st_as_sf()
gadm2 %>% glimpse()
```

```{r}
gadm2 %>% filter(NL_NAME_1 %in% c("埼玉県", "群馬県", "栃木県", "茨城県", "千葉県", "神奈川県", "東京都")) %>% 
  ggplot() + geom_sf(aes(fill = NAME_1)) + 
  ylim(34.7,37.2) + xlim(138.2,141) + 
  theme(legend.position = "none")
```



```{r}
library(geodata)
tokyo_prefecture <- gadm("JPN", level = 2, path = "Tokyo")
tokyo <- st_as_sf(tokyo_prefecture)
```
```{r}
tokyo %>% as_tibble()
```

```{r}
tokyo %>% colnames()
```

```{r}
tokyo %>% filter(NAME_1 == "Kanagawa") %>% 
  ggplot() + geom_sf(aes(fill = NAME_2)) + theme(legend.position = "none")
```

# Miscellaneous


```{r}
vietnam <- world_admin1 %>% filter(admin == "Vietnam")
```

```{r}
glimpse(vietnam)
```

```{r}
ggplot() +   geom_sf(data = vietnam) +   theme_minimal()
```

```{r}
world %>% arrange(region_wb) %>% pull(region_wb) %>% unique()
```

```{r}
southasia <- world %>% filter(region_wb == "South Asia")
ggplot() +   geom_sf(data = southasia) +   theme_minimal()
```

```{r}
southasia <- world %>% filter(region_wb == "Sub-Saharan Africa")
ggplot() +   geom_sf(data = southasia, aes(fill = admin)) +   theme(legend.position="none")
```


```{r}
world %>% arrange(region_wb) %>% pull(region_un) %>% unique()
```


```{r}
ggplot() +   geom_sf(data = world, aes(fill = admin)) +   theme(legend.position="none") +
  facet_wrap(.~region_wb)
```

```{r}
library(readxl)
gini_vietnam <- read_excel('le/DataVN.xls')
```

```{r}
glimpse(gini_vietnam)
```

```{r}
vietnam %>% arrange(name_alt) %>% distinct(name_alt)
```

```{r}
gini_vietnam %>% arrange(province) %>% distinct()
```

```{r}
vietnam %>% anti_join(gini_vietnam, by = c('name_alt' = 'province')) %>% arrange(name_alt) %>% distinct(name_alt)
```

```{r}
gini_vietnam %>% drop_na(gini) %>% anti_join(vietnam, by = c('province'= 'name_alt')) %>% arrange(province) %>% distinct(province)
```

```{r}
vietnam_gini <- vietnam %>% left_join(gini_vietnam, by = c("name_alt" = "province"))
```


```{r}
library(sf)
jpn <- st_read(system.file("extdata", "jpn.geojson", package = "sf"))
```

