---
title: "Map_Thesis"
author: Nguyen Thi Le
output: html_notebook
---

```{r}
getwd()
```

```{r}
library(tidyverse)
library(ggplot2)
library(sp)
library(raster)
library(rgdal)
library(viridis)
library(readxl)
library(maps)
library(geodata)
library("writexl")
library(rvest)
library(lubridate)
library(magrittr)
library (ggmap)
library(stringr)
```

# WORLD's MAP with country's Gini
## Get world map
```{r}
map.world <- map_data("world")
map.world

```

### One color version

```{r}

map.world %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "blue", size = 0.1)
```

### Colorful version

```{r}
p = ggplot(data=map.world, aes(x=long, y=lat, group=group, fill=factor(group))) 
p + geom_polygon(col="white") + theme(legend.position="none")

```
## Import GINI_world data

```{r}
giniw = read_excel('Giniworld.xlsx')
print(giniw)
```

## Join 2 datasets together
```{r}
map.world_joined <- left_join(map.world, giniw, by = c('region' = 'country'))
print(map.world_joined)
write.csv(map.world_joined,"map.world_joined.csv",row.names=TRUE)
```


```{r}
ggplot(data=map.world_joined, aes(x=long, y=lat, group=group)) + geom_polygon(aes(fill=Gini), show.legend = T) + scale_fill_gradientn(colours = c("lightblue", "darkblue")) + labs(x="Longitude", y="Latitude")
```


# Vietnam

## Get map data of Vietnam from GADM

OR
```{r}
#library(terra)
#mapvn0 <- gadm(country="VNM", level=1, path=tempdir())
#plot(mapvn0)
#mapvn0
#mapvn = as.data.frame(mapvn, row.names=NULL, optional=FALSE, geom=NULL)

```

OR
```{r}
library(sf)
library(tidyr)
mapvn0 <- geodata::gadm(country = "VNM", path = tempdir(), level = 1)
head(mapvn0)
plot(mapvn0)
mapvn <- st_as_sf(mapvn0) #the data was already saved as data frames after this command
head(mapvn)
write.csv(mapvn,"mapvn.csv",row.names=TRUE)
names(mapvn)

```

### Convert Geometry data into long/lat data???

I supposed that if I can convert geometry data into long/lat data, the map can be drawn.

I tried some solutions on the internet and the followings are one of them, but it did not work correctly.

Reference:
https://stackoverflow.com/questions/47661354/converting-geometry-to-longitude-latitude-coordinates-in-r
https://community.rstudio.com/t/convert-geometry-column-with-lists-of-longitude-latitude-coordinates/67114/7




### Import Vietnam dataset (Gini, GDP...)


```{r}
datavn = read_excel('DataVN.xls')
names(datavn)

```

### Join 2 dataset


```{r}
mapvn_joined <- left_join(mapvn, datavn, by = c('VARNAME_1' = 'province'))
mapvn_joined
summary(mapvn_joined)
names(mapvn_joined)
write.csv(mapvn_joined,"mapvn_joined.csv",row.names=TRUE)

#ggplot(data=mapvn_joined, aes(x=long, y=lat, group=group)) + geom_polygon(aes(fill=gini), show.legend = T) + scale_fill_gradientn(colours = c("lightblue", "darkblue")) + labs(x="Longitude", y="Latitude")
                

```
# Reference
https://rstudio-pubs-static.s3.amazonaws.com/647153_88833beacf2c4107ad76bc3bc7bc8b9d.html
https://rdrr.io/cran/terra/man/as.data.frame.html
This one: https://stackoverflow.com/questions/73825468/converting-spatvector-objects-to-data-frames-for-use-in-ggplot2





