---
title: "Responses to Assignment 3"
output: 
  html_notebook: 
    toc: yes
    toc_float: true
    highlight: haddock
    theme: cosmo
    number_sections: yes
---

## Assignment Three: The Week Four Assignment {-}

**WDI and `ggplot2`**

* Create an R Notebook of a Data Analysis containing the following and submit the rendered HTML file (eg. `a3_123456.nb.html`  by replacing 123456 with your ID)
  1. create an R Notebook using the R Notebook Template in Moodle,  save as `a3_123456.Rmd`, 
  2. write your name and ID and the contents, 
  3. run each code block, 
  4. preview to create `a3_123456.nb.html`,
  5. submit  `a3_123456.nb.html` to Moodle.

1. Choose at least one indicator of WDI

    - Information of the data: Name, Indicator, Description, Source, etc.
    - Download the data with `WDI`
    - Explain why you chose the indicator
    - List questions you want to study

2. Explore the data using visualization using `ggplot2`

    - Use a histogram (geom_histogram), boxplot (geom_boxplot), a scatter plot (geom_point), a line plot (geom_line)
    - For at least one chart, add title, and labels of axis, and add an explanation of it

3. Observations and difficulties encountered.

**Due:** 2023-01-16 23:59:00. Submit your R Notebook file in Moodle (The Third Assignment). Due on Monday!

# Set up

```{r}
library(tidyverse)
library(WDI)
```

It is not a must, but it is good to run the following code chunk and use `wdi_cache` for `WDIsearch()` to update the mata data. See `WDIcache` help. 

```{r cache = TRUE}
wdi_cache <- WDIcache()
```


# General Comments

## Varibles

We should know first about the variables. At least you must know if each of the variables is a categorical variable or a numerical variable.


## Example: 

## WDI

* IP.JRN.ARTC.SC: Scientific and technical journal articles
* FP.CPI.TOTL.ZG: Inflation, consumer prices (% annual growth) 
* NY.GDP.PCAP.KD: GDP per capita (constant 2015 US$)
* BX.KLT.DINV.CD.WD: Foreign direct investment, net inflows (BoP, current US$)
* SP.POP.TOTL (Total Population), MS.MIL.XPND.GD.ZS (Military expenditure (% of GDP)), MS.MIL.TOTL.P1 (Total Armed Forces Personnel), NY.GDP.MKTP.CD (NY.GDP.MKTP.CD: GDP)
* IP.JRN.ARTC.SC: Scientific and technical journal articles

# Responses to Questions

## Q1. 

```{r}
WDIsearch(string = "IP.JRN.ARTC.SC", field = "indicator", short = FALSE, cache = wdi_cache)
```

```{r cash=TRUE}
sct <- WDI(country = c("US", "FR", "KE"), indicator = c("IP.JRN.ARTC.SC"), start = 1990, end = NULL, extra = TRUE, cache = wdi_cache)
sct
```

```{r}
sct %>% summary()
```

```{r}
sct %>% filter(country == "France") %>%
  ggplot(aes(x = IP.JRN.ARTC.SC)) + ggtitle ("Distribution of number of scientific/technical articles in France") + 
  geom_histogram(bins = 8)
```

## Q2.

```{r}
WDIsearch(string = "gdp per capita", field = "name", short = FALSE)
```

```{r}
dat = WDI(indicator='NY.GDP.PCAP.KD', country=c('MX','CA','US'), start=2000, end=2020)
```
```{r}
ggplot(dat, aes(x = NY.GDP.PCAP.KD, fill=country), alpha = 0.5) +
  geom_histogram(bins = 10)
```


##

```{r}
FDI <- WDI(country = "all", indicator = "BX.KLT.DINV.CD.WD", start = 1960, end = NULL, extra = TRUE)
```

1. In the beginning, when I want to write_csv, I put the file in the file but not the data file. Then here came out an error which says can’t open the file;
2. When I use the same way of line chart and box plot chart want to draw the histogram, there is a mistake.

```{r}
write_csv(FDI, "data/FDI.csv")
```

```{r}
FDI %>% drop_na(BX.KLT.DINV.CD.WD) %>%
  ggplot(aes(BX.KLT.DINV.CD.WD)) + geom_histogram(bins = 10)
```


###

```{r}
artcpcap <- WDI(country = "all", indicator = "IP.JRN.ARTC.SC", start = 1960, end = NULL, extra = TRUE, cach = NULL) 
```

```{r}
artcpcap %>% 
  filter(income != "Aggregates", year == 2018, IP.JRN.ARTC.SC > 0, 
         income !="Not classified") %>%
  ggplot(aes(x = income, y = IP.JRN.ARTC.SC, fill = income)) + 
  geom_boxplot() + coord_trans(x ="identity", y="log10") 
  
```

```{r}
artcpcap %>% drop_na(IP.JRN.ARTC.SC) %>%
  filter(income !="Aggregates", IP.JRN.ARTC.SC > 0) %>%
  group_by(income, year) %>% summarize(median = median(IP.JRN.ARTC.SC)) %>%
  ggplot() + 
  geom_line(mapping = aes(x = year, y = median, color = income))
```


# Appendix: A secondary axis

* [Specify a secondary axis](https://ggplot2.tidyverse.org/reference/sec_axis.html)
  - This function is used in conjunction with a position scale to create a secondary axis, positioned opposite of the primary axis. All secondary axes must be based on a one-to-one transformation of the primary axes.

`scale_y_continuous(sec.axis = sec_axis(~ . scaling_function))`

## Example

Suppose you have two indicators, 

* NY.GDP.MKTP.KD: GDP (constant 2015 US$)
* NY.GDP.PCAP.KD: GDP per capita (constant 2015 US$)

```{r}
WDIsearch(string = "NY.GDP.MKTP.KD", field = "indicator", short = FALSE, cache = wdi_cache)
```

```{r}
WDIsearch(string = "NY.GDP.PCAP.KD", field = "indicator", short = FALSE, cache = wdi_cache)
```

List the name of countries of ASEAN and BRICs using `wdi_cache$country`.

```{r}
asean <- c("Brunei Darussalam", "Cambodia", "Lao PDR", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore")
brics <- c("Brazil", "Russian Federation", "India", "China")
```

Find the `iso2c` of the countries using `wdi_cache$country`.

```{r}
wdi_cache$country %>% 
  filter(country %in% 
           c("Brunei Darussalam", "Cambodia", "Lao PDR", "Myanmar", 
           "Philippines", "Indonesia", "Malaysia", "Singapore",
           "Brazil", "Russian Federation", "India", "China")) %>% 
  pull(iso2c)
```

Separate the iso3c's of the countries with commas and read data using `WDI`.

```{r cache=TRUE}
wdi_gdp <- WDI(
  country = c("BR", "BN", "CN", "ID", "IN", "KH", "LA", "MM", "MY", "PH", "RU", "SG"),
  indicator = c(gdp = "NY.GDP.MKTP.KD", gdpPercap = "NY.GDP.PCAP.KD"),
  start = 1960, extra = TRUE, cache = wdi_cache)
```


```{r}
wdi_gdp %>% filter(country %in% asean) %>% drop_na(gdp, gdpPercap) %>% summary() 
```

Using the summary, decide the scaling of two variables, `gdpPercap` and `gdp`.

```{r}
wdi_gdp %>% drop_na(gdp, gdpPercap) %>%
  filter(country %in% asean) %>%
  ggplot() + 
  geom_line(aes(x = year, y = gdpPercap, linetype = country)) + 
  geom_line(aes(x = year, y = gdp/(10^7), col = country)) + 
  coord_trans(x ="identity", y="log10") +
  scale_y_continuous(sec.axis = sec_axis(~ . *(10^7), name = "gdp/(10^7)"))
```

### BRICs

```{r}
wdi_gdp %>% filter(country %in% brics) %>% drop_na(gdp, gdpPercap) %>% summary() 
```

Using the summary, decide the scaling of two variables, `gdpPercap` and `gdp`.

```{r}
wdi_gdp %>% drop_na(gdp, gdpPercap) %>%
  filter(country %in% brics) %>%
  ggplot() + 
  geom_line(aes(x = year, y = gdpPercap, linetype = country)) + 
  geom_line(aes(x = year, y = gdp/(10^9), col = country)) + 
  coord_trans(x ="identity", y="log10") +
  scale_y_continuous(sec.axis = sec_axis(~ . *(10^7), name = "gdp/(10^9)"))
```

### test

```{r}
wdi_gdp %>% filter(country == "Brazil") %>% ggplot() + geom_histogram(aes(x = gdp))
```

```{r}
wdi_gdp %>% filter(country == "Brazil") %>% pull(gdp) %>% n()
```


```{r}
df4 <- WDI(
  country="all",
  indicator = c("SP.POP.TOTL","MS.MIL.XPND.GD.ZS","MS.MIL.TOTL.P1","NY.GDP.MKTP.CD"),
  start = 1960,
  end = NULL,
  extra = FALSE,
  cache = NULL,
  latest = NULL,
  language = "en"
)
```

```{r}
colnames(df4) <- c("Country","iso2c","iso3c","Year","Population","Mil_Exp","Total_Mil_HC","GDP")
df4
```

